---
title: "PP17"
author: "Sophia Chung"
format: html
editor: visual
---

# Question 1

a) $\sigma$: population SD, person-to-person variability in sleep hours for all CP students
b) prior SD of 0.5 hrs for $\mu$: uncertainty in population mean sleep hours for CP students, before observing data
c) SD of the Gamma prior distribution: uncertainty in population SD of sleep hours for CP students, before observing data
d) $\sigma/\sqrt{n}$: sample-to-sample variability in sample mean sleep hours over many hypothetical samples of CP students
e) posterior SD for $\mu$: uncertainty in population mean sleep hours for CP students, after observing data
f) posterior SD for $\sigma$: uncertainty in population SD of sleep hours for CP students, after observing data

# Question 2: Prior Predictive Distribution

- simulate $\mu$ from Normal(8, 0.5)
- independently simulation $\sigma$ from Gamma(16, 13)
- given $\mu$ and $\sigma$, simulate y from Normal($\mu$, $\sigma$)
- repeat many times and summarize y

# Question 3: Posterior Distribution

- simulate $\mu$ from Normal(8, 0.5)
- independently simulation $\sigma$ from Gamma(16, 13)
- given $\mu$ and $\sigma$, simulate a sample of 25 y values from Normal($\mu$, $\sigma$). for each sample, compute mean and SD
- repeat many times
- keep only repetitions for which the sample mean is 7.1 and the sample SD is 1.5
- summarize ($\mu$, $\sigma$)

# Question 4

```{r}
qnorm(c(0.01, 0.99), 7.32, 0.26)
```

There is posterior credibility of 98% that population mean sleep hours for CP students is between 6.7152 and 7.9249 hours.

# Question 5

```{r}
qnorm(c(0.01, 0.99), 1.46, 0.18)
```

There is posterior credibility of 98% that population SD sleep hours for CP students is between 1.0413 and 1.8787 hours.

# Question 6

No, since $\mu$ and $\sigma$ have a joint posterior distribution (Bivariate Normal with correlation 0.16). They are no longer independent like they were under the prior distribution.

# Question 7: Posterior Predictive

- simulate ($\mu$, $\sigma$) from the joint posterior distribution
- given ($\mu$, $\sigma$), simulate a value of y from Normal(7.32, 1.46)
- repeat many times and summarize y
- find the 2.5th and 97.5th percentiles for the endpoints of a 95% posterior prediction interval

# Question 8

There is a 95% posterior predictive probability that the sleep hours for a randomly selected CP student are between 4.42 and 10.32.

# Question 9

- simulate ($\mu$, $\sigma$) from the joint posterior distribution and compute $\sigma/\mu$
- repeat many times and summarize $\sigma/\mu$
- find the 1st and 99th percentiles for the endpoints of a 98% posterior credible interval