---
title: "PP20"
author: "Sophia Chung"
format: html
editor: visual
---

# Question 1

Regardless of whether we think there might be a difference or not, our beliefs about $\mu_1$ influence our beliefs about $\mu_2$.

# Question 2: Posterior of $\mu_1$ and $\mu_2$

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
library(janitor)
```

```{r}
mu0 <- 37
tau0 <- 20

sigma <- 10
n <- 116
ybar <- 30.26

tau_n <- 1/sqrt(n/sigma^2+1/tau0^2)
mu_n <- mu0*(1/tau0^2)/(1/tau_n^2)+ybar*(n/sigma^2)/(1/tau_n^2)

df <- data.frame(c("Precision", "SD", "Mean"),
            c(1/tau0^2, tau0, mu0),
            c(n/sigma^2, sigma/sqrt(n), ybar),
            c(1/tau_n^2, tau_n, mu_n))

kable(df, digits=3, align='r',
  col.names=c("", "Prior", "Data (Sample Mean)", "Posterior"))
```

```{r}
mu1_mean <- mu_n
mu1_sd <- tau_n
c(mu1_mean, mu1_sd)
```

The posterior distribution of $\mu_1$ is Normal(30.27, 0.93).

```{r}
mu0 <- 37
tau0 <- 20

sigma <- 10
n <- 112
ybar <- 16.31

tau_n <- 1/sqrt(n/sigma^2+1/tau0^2)
mu_n <- mu0*(1/tau0^2)/(1/tau_n^2)+ybar*(n/sigma^2)/(1/tau_n^2)

df <- data.frame(c("Precision", "SD", "Mean"),
            c(1/tau0^2, tau0, mu0),
            c(n/sigma^2, sigma/sqrt(n), ybar),
            c(1/tau_n^2, tau_n, mu_n))

kable(df, digits=3, align='r',
  col.names=c("", "Prior", "Data (Sample Mean)", "Posterior"))
```

```{r}
mu2_mean <- mu_n
mu2_sd <- tau_n
c(mu2_mean, mu2_sd)
```

The posterior distribution of $\mu_2$ is Normal(16.36, 0.94).

# Question 3

```{r}
mu_diff_mean <- mu1_mean-mu2_mean
mu_diff_sd <- sqrt(mu1_sd^2+mu2_sd^2)
c(mu_diff_mean, mu_diff_sd)
```

The posterior distribution of $\mu_1 - \mu_2$ is Normal(13.92, 1.32). 

# Question 4

```{r}
qnorm(c(0.25, 0.75), mu_diff_mean, mu_diff_sd)
```

There is a posterior probability of 50% that mean response to the UN Africa question is between 13.03 and 14.81 percentage points greater for those who see X = 65 than X = 10.

```{r}
qnorm(c(0.1, 0.9), mu_diff_mean, mu_diff_sd)
```

There is a posterior probability of 80% that mean response to the UN Africa question is between 12.22 and 15.61 percentage points greater for those who see X = 65 than X = 10.

```{r}
qnorm(c(0.01, 0.99), mu_diff_mean, mu_diff_sd)
```

There is a posterior probability of 98% that mean response to the UN Africa question is between 10.84 and 17 percentage points greater for those who see X = 65 than X = 10.

# Question 5

```{r}
1 - pnorm(0, mu_diff_mean, mu_diff_sd)
```

After observing data, there is 100% posterior probability that $\mu_1 > \mu_2$.

# Question 6

```{r}
qnorm(c(0.25, 0.75), mu_diff_mean, mu_diff_sd)/sigma
```

There is a posterior probability of 50% that mean response to the UN Africa question is between 1.3 and 1.48 standard deviations greater for those who see X = 65 than X = 10.

```{r}
qnorm(c(0.1, 0.9), mu_diff_mean, mu_diff_sd)/sigma
```

There is a posterior probability of 80% that mean response to the UN Africa question is between 1.22 and 1.56 standard deviations greater for those who see X = 65 than X = 10.

```{r}
qnorm(c(0.01, 0.99), mu_diff_mean, mu_diff_sd)/sigma
```

There is a posterior probability of 98% that mean response to the UN Africa question is between 1.08 and 1.7 standard deviations greater for those who see X = 65 than X = 10.

# Question 7

Yes, all of the credibility in our posterior distribution is on $\mu_1 > \mu_2$.

# Question 8

Yes, our posterior distribution implies that there is a large effect size.

# Question 9

The study randomly assigns the anchor values, so yes, there is evidence that the difference in mean responses is caused by the anchor.